---
output: github_document
bibliography: vignettes/references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# massProps

<!-- badges: start -->
<!-- badges: end -->

The `massProps` package extends `rollupTree` with functions to recursively calculate mass properties (and optionally, their uncertainties) for arbitrary composition trees. Formulas implemented are described in a technical paper published by the Society of Allied Weight Engineers.[@zimmerman:05:sawe]

## Installation

```{r eval = FALSE}
install.packages("massProps")
```

You can install the development version of massProps from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jsjuni/massProps")
```

## Example

```{r echo = FALSE}
suppressPackageStartupMessages({library(igraph)})
```

Suppose we have the following mass properties table:

```{r}
library(massProps)
test_table
```

and this graph with edges representing child-parent relations:

```{r}
library(igraph)
E(test_tree)
```
```{r echo = FALSE}
plot(test_tree,layout=layout_as_tree(test_tree, 2, mode="in"), vertex.shape = 'none', edge.arrow.mode = 0)
```

We can roll up mass properties to non-leaf elements as follows:

```{r}
rollup_mass_props(test_tree, test_table)
```

The input data may include mass properties uncertainties as well:

```{r}
test_unc_table
```

Mass properties and their uncertainties can be rolled up as follows:

```{r}
rollup_mass_props_and_unc(test_tree, test_unc_table)
```

## References
