---
title: "massProps"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{massProps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(massProps)
```

# Overview

The `massProps` package extends `rollupTree` with functions to recursively calculate mass properties (and optionally, their uncertainties). Formulas implemented are described in a technical paper published by the Society of Allied Weight Engineers.[@sawe:05]

# Synopsis

## Data Structures

`massProps` operates on two fundamental data structures: a mass properties table and a tree. The mass properties table has an entry for every item in a tree structure of items; the edges of the tree convey the parent-child relations among items. The two data structures are linked by the `id` column of the data frame, which must be a character vector of unique item identifiers, and the vertex names of the tree. The sets of identifiers must be identical.

### Mass Property Table

#### Required Columns for Mass Properties

The Mass Property Table must contain the following columns. Other columns may exist and will remain unmodified.

-   `id` unique identifier for each item (row)

-   `mass` mass of the item (numeric)

-   `Cx` $x$-component of center of mass (numeric)

-   `Cy` $y$-component of center of mass (numeric)

-   `Cx` $z$-component of center of mass (numeric)

-   `Ixx` moment of inertia about the $x$ axis (numeric)

-   `Iyy` moment of inertia about the $y$ axis (numeric)

-   `Izz` moment of inertia about the $z$ axis (numeric)

-   `Ixy` product of inertia relative to the $x$ and $y$ axes (numeric)

-   `Ixz` product of inertia relative to the $x$ and $z$ axes (numeric)

-   `Iyz` product of inertia relative to the $y$ and $z$ axes (numeric)

-   `POIconv` either '+' or '-', indicating the sign convention for products of inertia[^1]

-   `Ipoint` logical indicator that this item is considered a point mass (i.e., its inertia contribution is negligible)[^2]

[^1]: In the negative convention, for example, $I_{XY} \equiv -\int xy \rho \thinspace dV$. In the positive convention, $I_{XY} \equiv \int xy \rho \thinspace dV$.

[^2]: The same algebraic result can be achieved by setting all moments and products of inertia to zero, but `rollup_mass_props()` by default ensures that all leaf items in the tree have mass properties that correspond to physically-realizable objects. A zero inertia tensor will fail this check. Rather than relax the check (which is essential for trustworthy results), a `TRUE` value for `Ipoint` indicates that the inertia tensor should be excluded from computations. \`\`

#### Required Columns for Mass Properties Uncertainty

The following columns are required for uncertainty calculations:

-   `σ_mass` mass uncertainty (numeric)

-   `σ_Cx` $x$-component of center of mass uncertainty (numeric)

-   `σ_Cy` $y$-component of center of mass uncertainty (numeric)

-   `σ_Cx` $z$-component of center of mass uncertainty (numeric)

-   `σ_Ixx` moment of inertia about the $x$ axis uncertainty (numeric)

-   `σ_Iyy` moment of inertia about the $y$ axis uncertainty (numeric)

-   `σ_Izz` moment of inertia about the $z$ axis uncertainty (numeric)

-   `σ_Ixy` product of inertia relative to the $x$ and $y$ axes uncertainty (numeric)

-   `σ_Ixz` product of inertia relative to the $x$ and $z$ axes uncertainty (numeric)

-   `σ_Iyz` product of inertia relative to the $y$ and $z$ axes uncertainty (numeric)

It is the caller's responsibility to ensure that all values are expressed in appropriate and compatible units.

#### Tree

The tree is an `igraph::graph` with vertices named by identifiers in the mass properties table. The graph must satisfy certain well-formedness properties:

-   it is connected an acyclic (as an undirected graph), i.e., it is a tree

-   it is directed, with edge direction going from child to parent

-   it contains neither loops (self-edges) or multiple edges

-   it contains a single root vertex (i.e., one whose out degree is zero)

## Invocation

Suppose we have the following mass properties table:

```{r}
test_table
```

# Theory

### Mass

$$
\newcommand{\sumwi}{\sum_{i=1}^{n}w_i}
W = \sumwi
$$

### Center of Mass

$$
\newcommand{\sumwivi}[1]{\sum_{i=1}^{n}{w_i}{{#1}_i}}
\begin{align}
\bar{x} & = \sumwivi{x} \bigg/ \sumwi \\
\bar{y} & = \sumwivi{y} \bigg/ \sumwi \\
\bar{z} & = \sumwivi{z} \bigg/ \sumwi \\
\end{align}
$$

## Inertia Tensor

##### Moments of Inertia

$$
\newcommand{\moia}[3]{
\sum_{i=1}^{n} \left[ {I_{#1}}_i
  + w_i \left( {#2}_i^2 + {#3}_i^2 \right)
  - w_i \left( \bar{#2}^2 + \bar{#3}^2 \right)
\right]
}
\newcommand{\moib}[3]{
\sum_{i=1}^{n} \left\{ {I_{#1}}_i
  + w_i \left[ \left( {#2}_i - \bar{#2} \right)^2 + \left( {#3}_i - \bar{#3} \right)^2 \right]
\right\}
}
\begin{align}
I_{XX} & = \moia{XX}{y}{z} & = \moib{XX}{y}{z} \\
I_{YY} & = \moia{YY}{x}{z} & = \moib{YY}{x}{z} \\
I_{ZZ} & = \moia{ZZ}{x}{y} & = \moib{ZZ}{x}{y} \\
\end{align}
$$

#### Products of Inertia

#### $$
\newcommand{\di}[1]{({#1}_i - \bar{#1})}
\newcommand{\poia}[3]{\sum_{i=1}^{n} \left[ {I_{#1}}_i + w_i {{#2}_i}{{#3}_i} -w_i (\bar{#2}\bar{#3})\right]}
\newcommand{\poib}[3]{\sum_{i=1}^{n} \left[ {I_{#1}}_i + w_i \di{#2}\di{#3}\right]}
\begin{align}
I_{XY} & = \poia{XY}{x}{y} & = \poib{XY}{x}{y} \\
I_{XZ} & = \poia{XZ}{x}{z} & = \poib{XZ}{x}{z} \\
I_{YZ} & = \poia{YZ}{y}{z} & = \poib{YZ}{y}{z} \\
\end{align}
$$ Matrix Formulation

Let

$$
\begin{align}
\boldsymbol{d}_i & = (\di{x} \quad \di{y} \quad \di{z})^T \\
\boldsymbol{Q}_i & = \boldsymbol{d}_i {\boldsymbol{d}_i}^T
\end{align}
$$ Then

$$
\begin{align}
\boldsymbol{Q}_i & =
\begin{bmatrix}
\di{x}^2     & \di{x}\di{y} & \di{x}\di{z} \\
\di{y}\di{x} & \di{y}^2     & \di{y}\di{z} \\
\di{z}\di{x} & \di{z}\di{y} & \di{z}^2     \\
\end{bmatrix}
\end{align}
$$ Let $\boldsymbol{s}_i$ be the matrix of inertia tensor summands in the standard formulas for a given subcomponent $i$ above. Note that the products of inertia formulas clearly employ the positive sign convention. Consequently, we negate those values in the inertia tensor. $$
\begin{align}
\boldsymbol{s}_i & = \boldsymbol{I}_i \\
& -
w_i \begin{bmatrix}
-\di{y}^2 - \di{z}^2 & \di{x}\di{y}         & \di{x}\di{z}         \\
\di{x}\di{y}         & -\di{x}^2 - \di{z}^2 &  \di{y}\di{z}        \\
\di{x}\di{z}         & \di{y}\di{z}         & -\di{x}^2 - \di{y}^2 \\
\end{bmatrix} \\
& =
\boldsymbol{I}_i \\
& -
w_i \begin{bmatrix}
\di{x}^2     & \di{x}\di{y} & \di{x}\di{z} \\
\di{x}\di{y} & \di{y}^2     & \di{y}\di{z} \\
\di{x}\di{z} & \di{y}\di{z} & \di{z}^2     \\
\end{bmatrix} \\
& -
w_i \begin{bmatrix}
-\di{x}^2 - \di{y}^2 - \di{z}^2 & 0                               & 0                              \\
0                               & -\di{x}^2 - \di{y}^2 - \di{z}^2 & 0                              \\
0                               & 0                               &-\di{x}^2 - \di{y}^2 - \di{z}^2 \\
\end{bmatrix} \\
& = 
\boldsymbol{I}_i
- 
w_i \left[ \boldsymbol{Q}_i - \mathrm{tr}(\boldsymbol{Q}_i) \boldsymbol{I}_3 \right]
\end{align}
$$

## Mass Property Uncertainties

### Mass Uncertainty

$$
\sigma_w = \sqrt{ \sum_{i=1}^n {{\sigma_w}_i}^2 }
$$

### Center of Mass Uncertainty

$$
\newcommand{\sigmacm}[1]{
\sqrt{ \sum_{i=1}^n \big\{ (w_i {\sigma_{#1}}_i)^2 + [{\sigma_w}_i \di{#1}]^2 \big\}\bigg/ \sum_{i=1}^n w_i}
}
\begin{align}
\sigma_x & = \sigmacm{x} \\
\sigma_y & = \sigmacm{y} \\
\sigma_z & = \sigmacm{z} \\
\end{align}
$$

### Inertia Tensor Uncertainty

#### Moments of Inertia Uncertainties

$$
\newcommand{\sigmamoi}[3]{\sqrt{ \sum_{i=1}^n \big\{  \sigma_{{I_{#1}}_i}^2 + \big[ 2 w_i \di{#2} \sigma_{{#2}_i} \big]^2 + \big[ 2 w_i \di{#3} \sigma_{{#3}_i} \big]^2 + \big[ \big(\di{#2}^2 + \di{#3}^2 \big)\sigma_{w_i}\big]^2 \big\} }}
\begin{align}
\sigma_{I_{XX}} & = \sigmamoi{XX}{y}{z} \\
\sigma_{I_{YY}} & = \sigmamoi{YY}{x}{z} \\
\sigma_{I_{ZZ}} & = \sigmamoi{ZZ}{x}{y} \\
\end{align}
$$

#### Products of Inertia Uncertainties

$$
\begin{align}
\newcommand{\sigmapoi}[3]{\sqrt{ \sum_{i=1}^n \big\{  \sigma_{{I_{#1}}_i}^2 + \big[ \di{#2} w_i \sigma_{{#3}_i} \big]^2 + \big[  \di{#2}\di{#3}\sigma_{w_i} \big]^2 + \big[ \di{#3}  w_i \sigma_{{#2}_i} \big]^2 \big\} }}
\sigma_{I_{XY}} & = \sigmapoi{XY}{x}{y} \\
\sigma_{I_{XZ}} & = \sigmapoi{XZ}{x}{z} \\
\sigma_{I_{YZ}} & = \sigmapoi{YZ}{y}{z} \\
\end{align}
$$

#### Matrix Formulation

Let

$$
\begin{align}
\boldsymbol{d}_i & = (\di{x} \quad \di{y} \quad \di{z})^T \\
{\boldsymbol{\sigma}_{c_m}}_i & = (\sigma_{x_i} \quad \sigma_{y_i} \quad \sigma_{z_i})^T \\
\boldsymbol{P}_i & = \boldsymbol{d}_i {\boldsymbol{\sigma}_{c_m}}_i^T \\
\boldsymbol{Q}_i & = \boldsymbol{d}_i {\boldsymbol{d}_i}^T
\end{align}
$$

Then

$$
\begin{align}
\boldsymbol{P}_i & = 
\begin{bmatrix}
(x_i - \bar{x})\sigma_{x_i} &\di{x}\sigma_{y_i} &\di{x}\sigma_{z_i} \\
\di{y}\sigma_{x_i} & \di{y}\sigma_{y_i} & \di{y}\sigma_{z_i} \\
\di{z}\sigma_{x_i} & \di{z}\sigma_{y_i} & \di{z}\sigma_{z_i} \\
\end{bmatrix}
\\
\\
\boldsymbol{Q}_i & =
\begin{bmatrix}
(x_i - \bar{x})^2 &\di{x}\di{y} &\di{x}\di{z} \\
\di{y}(x_i - \bar{x}) & \di{y}^2 & \di{y}\di{z} \\
\di{z}(x_i - \bar{x}) & \di{z}\di{y} & \di{z}^2 \\
\end{bmatrix}
\end{align}
$$

Let $\boldsymbol{s}_i^2$ be the matrix of inertia tensor uncertainty summands in the standard formulas for a given subcomponent $i$ above. Let ${p_X}_i$, ${p_Y}_i$, and ${p_Z}_i$ be the respective diagonal elements of $P_i$. If we interpret squaring a matrix as the Hadamard (element-wise) product with itself, then $$
\begin{align}
\boldsymbol{s}_i^2 & = {\boldsymbol{\sigma_I}}_i^2 \\
& + 
\begin{bmatrix}
2 w_i \di{y} \sigma_{y_i} &   w_i\di{x} \sigma_{y_i} &   w_i\di{x} \sigma_{z_i} \\
  w_i\di{x} \sigma_{y_i} & 2 w_i\di{x} \sigma_{x_i} &   w_i \di{y} \sigma_{z_i} \\
  w_i\di{x} \sigma_{z_i} &   w_i \di{y} \sigma_{z_i} & 2 w_i\di{x} \sigma_{x_i}
\end{bmatrix}^2 \\
& +
\begin{bmatrix}
2 w_i \di{z} \sigma_{z_i} &   w_i \di{y} \sigma_{x_i} &   w_i \di{z} \sigma_{x_i} \\
  w_i \di{y} \sigma_{x_i} & 2 w_i \di{z} \sigma_{z_i} &   w_i \di{z} \sigma_{y_i} \\
  w_i \di{z} \sigma_{x_i} &   w_i \di{z} \sigma_{y_i} & 2 w_i \di{y} \sigma_{y_i}
\end{bmatrix}^2 \\
& +
\begin{bmatrix}
(\di{y}^2 + \di{z}^2)\sigma_{w_i} &\di{x}\di{y}\sigma_{w_i} &\di{x}\di{z}\sigma_{w_i} \\
\di{y}(x_i - \bar{x})\sigma_{w_i} & ((x_i - \bar{x})^2 + \di{z}^2)\sigma_{w_i} & \di{y}\di{z}\sigma_{w_i} \\
\di{z}(x_i - \bar{x})\sigma_{w_i} & \di{z}\di{y}\sigma_{w_i} & ((x_i - \bar{x})^2 + \di{y}^2)\sigma_{w_i} \\
\end{bmatrix}^2 \\ \\
& = {\boldsymbol{\sigma_I}}_i^2 \\
& + w_i^2 \left(
\boldsymbol{P}_i -
\begin{bmatrix}
(x_i - \bar{x})\sigma_{x_i} + 2 \di{y}\sigma_{y_i} & 0 & 0 \\
0 & \di{y}\sigma_{y_i} + 2\di{x}\sigma_{x_i} & 0 \\
0 & 0 & \di{z}\sigma_{y_i} + 2\di{x}\sigma_{x_i} \\
\end{bmatrix}
\right) ^2 \\
& + w_i^2 \left(
\boldsymbol{P}_i^T -
\begin{bmatrix}
(x_i - \bar{x})\sigma_{x_i} + 2 \di{z}\sigma_{y_i} & 0 & 0 \\
0 & \di{y}\sigma_{y_i} + 2 \di{z}\sigma_{z_i} & 0 \\
0 & 0 & \di{z}\sigma_{y_i} + 2 \di{y}\sigma_{y_i} \\
\end{bmatrix}
\right) ^2 \\
& + \sigma_{w_i}^2 \left(
\boldsymbol{Q}_i - \mathrm{tr}(\boldsymbol{Q}_i)\boldsymbol{I}_3
\right)^2 \\ \\
& = {\boldsymbol{\sigma_I}}_i^2 \\
& + w_i^2 \left(
\boldsymbol{P}_i -
\begin{bmatrix}
 {p_X}_i + 2 {p_Y}_i & 0 & 0 \\
0 & {p_Y}_i + 2 {p_X}_i & 0 \\
0 & 0 & {p_Z}_i + 2  {p_X}_i \\
\end{bmatrix}
\right) ^2 \\
& + w_i^2 \left(
\boldsymbol{P}_i^T -
\begin{bmatrix}
 {p_X}_i + 2{p_Z}_i & 0 & 0 \\
0 & {p_Y}_i + 2 {p_Z}_i & 0 \\
0 & 0 & {p_Z}_i + 2 {p_Y}_i \\
\end{bmatrix}
\right) ^2 \\
& + \sigma_{w_i}^2 \left(
\boldsymbol{Q}_i - \mathrm{tr}(\boldsymbol{Q}_i)\boldsymbol{I}_3
\right)^2
\end{align}
$$
